Effect of quality-adjusted production function inputs

### 2015-06-24 Articles/Presentation_ESEE_2015/ESEE_2015.Rnw
Adding data to CST models.

<<Fitted_and_Residuals>>=
fittedAndResidData <- function(model, id){
  df <- data.frame(select(getData(model), Year))
  # Trim fitted to be same length as df to eliminate problems with possible NAs at end of fitted
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  df <- na.omit(df)
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}

temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  return(fittedAndResidData(model, id))  
}) %>% do.call(rbind, .)
temp_data$cst <- FALSE

# temp_data_CST <- leaf_apply(modelsCST,class=c("plm", "lm"), f = function(model, id){
#   return(fittedAndResidData(model, id))
# }) %>% do.call(rbind, .)
# temp_data_CST$cst <- TRUE

# Get the unadjusted, without-energy residuals and store them in a data frame
r_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
               select(Year, Country, Residuals, model)
colnames(r_unadj_noE)[colnames(r_unadj_noE)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
fitted_and_resid_data <- merge(temp_data, r_unadj_noE)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

@@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

@@

Figures~\ref{fig:Graphs CD-resid}--\ref{fig:Graphs CESekl-resid}
show fitting residuals for all models.
%
<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD"))
@
%
<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest == "lek" | nest=="kl")))
@
%
<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & (nest=="ekl" | nest=="kl")))
@

%++++++++++++++++++++++++++++++
\subsection{Dynamic Solow residuals} 
\label{sec:dynamic_solow_residuals}
%++++++++++++++++++++++++++++++

<<Dynamic Solow Residuals>>=
# Calculate finite difference approximations of time derivatives for output and all factors of production
dynamicSR <- function(model, id, lag=1, pad="tail"){
  id_list <- strsplit(id, split = "[.]")[[1]]
  energy <- id_list[[3]]
  modelStr <- id_list[[4]]

  f <- model$formula
  denom <- summands(f, n=length(summands(f))) # time variable
  nums <- all.vars(f) [1:(length(all.vars(f))-1)] # y, k, l, and e (if present)
  df <- getData(model)[c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)

  # Calculate names for the columns of finite difference derivatives
  derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], ".d", denom)}) %>% unlist
  derivCols <- lapply(1:length(nums), function(i){
    # Calculate finite difference derivative columns
    ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
  }) %>% do.call(cbind, .) %>% as.data.frame
  names(derivCols) <- derivNames  
  df <- cbind(derivCols, df[ , c(nums, denom)]) # puts derivatives first, then raw numbers, then time
  df[["dydtOvery"]] <- 1 / df[[ nums[[1]] ]] * df[[ derivNames[[1]] ]]
  coeffs <- naturalCoef(model)
  df <- cbind(df, coeffs)

  if (modelStr == "CD"){
    lambda_fd <- 
      df[["dydtOvery"]] - 
      df[["alpha_1"]] / df[[nums[[2]]]] * df[[derivNames[[2]]]] - 
      df[["alpha_2"]] / df[[nums[[3]]]] * df[[derivNames[[3]]]]
    if (energy == "withE"){
      # Subtract the energy term also
      lambda_fd <- lambda_fd - 
        df[["alpha_3"]] / df[[nums[[4]]]] * df[[derivNames[[4]]]]
    }
    # Fraction of growth explained by Solow Residual
    solowFrac <- lambda_fd / df[["dydtOvery"]]
    # Add to data frame
    df$lambda_fd <- lambda_fd
    df$solowFrac <- solowFrac
    return(df)
  } else if (modelStr == "CES"){
    # Assign nice variable names so that the equations below are easier to read (and debug, and maintain)
    # This code is nest-aware.
    n <- model$nest
    for (i in 1:length(model$nest)){
      df[[paste0("x", i)]] <- df[[ nums[ model$nest[i]+1 ] ]] # GDP is 1st col, add 1 to get to the factors of production
      df[[paste0("dx", i, "dt")]] <- df[[ derivNames[ model$nest[i]+1 ] ]]
    }
    if (energy == "noE"){
      # Calculate terms for the finite-difference Solow residual (lambda_fd)
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else if (energy == "withE"){
      df <- df %>% 
        transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
        transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
        transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
        transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
        transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
        transform(solowFrac = lambda_fd / dydtOvery)
      return(df)
    } else {
      stop (paste("Unknown energy:", energy))
    }
  } else {
    stop(paste("Unknown modelStr", modelStr))
  }
}

lag.to.use <- 1 # years
temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
                        f = function(model, id, lag=lag.to.use, pad="tail"){
  df <- dynamicSR(model, id=id, lag=lag, pad=pad)
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  df$Year <- getData(model)[["Year"]]
  return(df)  
}) %>% do.call(plyr::rbind.fill, .)

# Get the unadjusted, without-energy Solow residuals and store them in a data frame
sr_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
                select(Year, Country, lambda_fd, model)
colnames(sr_unadj_noE)[colnames(sr_unadj_noE)=="lambda_fd"] <- "sr_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
solowResiduals <- merge(temp_data, sr_unadj_noE)

# Create a column containing ∆lambda_df, the difference between a lambda_df value
# and the lambda_df value in the same year for the unadjusted, no-energy case
solowResiduals[["Dlambda_fd"]] <- abs(solowResiduals[["lambda_fd"]]) - abs(solowResiduals[["sr_unadj_noE"]])

# Relevel factors before plotting
solowResiduals$Country <- relevelFactor(as.factor(solowResiduals$Country), c("PT", "UK"))
solowResiduals$flavor <- relevelFactor(as.factor(solowResiduals$flavor), c("Unadjusted", "Quality-adjusted"))
solowResiduals$energy <- relevelFactor(as.factor(solowResiduals$energy), c("Without energy", "With energy"))
solowResiduals$nest <- relevelFactor(as.factor(solowResiduals$nest), c("kl", "kle", "lek", "ekl"))

# Make data viewing easier
solowResiduals <- arrange(solowResiduals, Country, model, flavor, energy, nest)
@

<<DynamicSR_Graph>>=
dynamicSRGraph <- function(data){
ggplot(data=data) +
  geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=dydtOvery)) +
  geom_line(mapping=aes(x=Year, y=lambda_fd, linetype=energy)) +
  scale_linetype_manual(values=c(2,3), name="") +
  facet_grid(Country ~ flavor) +
  scale_x_continuous(breaks=c(1960, 1980, 2000)) +
  xlab("") + 
  ylab(expression(lambda[ij])) +
  xy_theme()  
}
@

``Dynamic'' Solow residuals~($\lambda_{i,j}$) can be calculated 
for the Cobb-Douglas and CES models
as shown in Appendices~\ref{sec:dynamic_sr_CD} and~\ref{sec:dynamic_sr_CES},
respectively.

Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
of the Solow residual~($\lambda_{i,j}$) over time
as calculated by Equation~\ref{eq:lambdaij} 
with the timespan for finite differences ($t_j - t_i$) set to 
\Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
%
<<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CD"))
@
%
<<DynamicSR_CESkle_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($kl$)$e$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="kle" | nest=="kl")))
@
%
<<DynamicSR_CESlek_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($le$)$k$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="lek" | nest=="kl")))
@
%
<<DynamicSR_CESekl_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the CES model with ($ek$)$l$ nesting. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
dynamicSRGraph(filter(solowResiduals, model=="CES" & (nest=="ekl" | nest=="kl")))
@

@@

  scale_shape(solid=FALSE) +
  guides(shape=FALSE) +
  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=Dlambda, color=energy, shape=cst)) +

@@

  scale_shape(solid=FALSE) +
  guides(shape=FALSE) +
  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=percDlambda, color=energy, shape=cst)) +

@@

ggplot(data=filter(AllCoeffs, model=="CES")) +
  geom_point(mapping=aes(x=Dmse, y=Dlambda, color=energy, shape=nest)) +

@@

ggplot(data=filter(AllCoeffs, model=="CES")) +
  geom_point(mapping=aes(x=Dmse, y=percDlambda, color=energy, shape=nest)) +

@@

%++++++++++++++++++++++++++++++
\subsection{Does quality-adjusting the factors of production 
            decrease fitting residuals and/or the dynamic Solow residual?} 
\label{sec:quality_adj_energy_and_dynamic_solow_residual}
%++++++++++++++++++++++++++++++

<<SR and R trajectory data>>=
traj.data <- merge(solowResiduals[c("Country", "Year", "model", "flavor", "energy", "nest", "Dlambda_fd")], 
                   fitted_and_resid_data[c("Country", "Year", "model", "flavor", "energy", "nest", "Dr")]) %>%
                   arrange(Country, model, flavor, energy, nest)
traj.data.stats <- traj.data %>% 
  group_by(model, Country, flavor, energy, nest) %>%
  summarise(Dlambda_fd.mean = mean(Dlambda_fd, na.rm=TRUE), Dr.mean = mean(Dr))

############################################
# Dlambda_fd.p <- traj.data %>% 
#   group_by(model, Country, flavor, energy, nest) %>% 
#   select(Dlambda_fd) %>%
#   t.test(., alternative="less")[["p.value"]]
############################################  
@

<<SR and R trajectory plot function>>=
residualsTrajGraph <- function(data){
  plot <- ggplot(data=data) + 
    geom_point(mapping=aes(x=Dr, y=Dlambda_fd, color=energy)) + 
    facet_grid(Country ~ flavor) +
    xlab(expression(Delta*r[i])) +
    ylab(expression(Delta*lambda[ij])) +
    xy_theme() + 
    theme(legend.title=element_blank())
  return(plot)
}
@

We can test our hypothesis that quality-adjusting the factors of production
and including energy will reduce the Solow residuals~($\lambda$)
and the fitting residuals~($r$) by plotting
$\Delta \lambda_i$ vs. $\Delta r_i$, where
%
\begin{equation} \label{eq:delta_lambda_i}
  \Delta \lambda_i \equiv \left| \lambda_i \right| - \left| \lambda_{i, \, Unadjusted, \, Without \, energy} \right|
\end{equation}
%
and
%
\begin{equation} \label{eq:delta_r}
  \Delta r_i \equiv \left| r_i \right| - \left| r_{i, \,Unadjusted, \, Without \, energy} \right|.
\end{equation}
%
If most of the points fall in the lower-left quadrant of a graph, 
we can conclude that both Solow residual~($\lambda$) and
fitting residual~($r$) have decreased due to either 
quality-adjustment of the factors of production or 
including energy.

Figure~\ref{fig:Solow_residual_trajectories_CD} indicates that
for the Cobb-Douglas model, 
neither 
quality-adjusting the factors of production
nor including energy 
consistently reduces
the Solow residual or the fitting residual.

<<Solow_residual_trajectories_CD, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the Cobb-Douglas model.">>=
residualsTrajGraph(filter(traj.data, model=="CD"))
@

We can also calculate the average values for $\Delta \lambda_{i,j}$ and $\Delta r_{i}$ 
as shown in Table~\ref{tab:Trajectory_Means_CD}.
%
<<Trajectory_Means_CD, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CD") %>% select(-nest), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for Cobb-Douglas models.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@
%
The following tables and figures show the same analysis for the CES model.
%
<<Solow_residual_trajectories_CESkle, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($kl$)$e$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="kle")))
@
%
<<Trajectory_Means_CESkle, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="kle")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($kl$)$e$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

<<Solow_residual_trajectories_CESlek, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($le$)$k$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="lek")))
@

<<Trajectory_Means_CESlek, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="lek")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($le$)$k$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

<<Solow_residual_trajectories_CESekl, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residual and fitting residual trajectories relative to the Unadjusted, Without-energy condition for the CES model with ($ek$)$l$ nesting.">>=
residualsTrajGraph(filter(traj.data, model=="CES" & (nest=="kl" | nest=="ekl")))
@

<<Trajectory_Means_CESekl, results="asis">>=
myXTable(data = filter(traj.data.stats, model=="CES" & (nest=="kl" | nest=="ekl")), 
         caption = "Average $\\Delta\\lambda$ and $\\Delta r$ for CES models with ($ek$)$l$ nesting.",
         label=paste0("tab:", opts_current$get("label")),
         digits=6)
@

@@

>>>>

<<Fitted_and_Residuals>>=
fittedAndResidData <- function(model, id){
  df <- data.frame(select(model$data, Year)) # getData doesn't work for lm models that are created for CST=TRUE.
  # By including [c(1:length(df$Year))] on the end, we 
  # add NAs (as appropriate) to the end of these columns
  # if data do not exist.
  df$fitted <- yhat(model)[c(1:length(df$Year))]
  df$Residuals <- resid(model)[c(1:length(df$Year))]
  # Now add the response, using the original variable name.
  df[[summands(model$formula, right = FALSE)]] <- response(model)[c(1:length(df$Year))]
  # Now add meta information
  id_list <- strsplit(id, split = "[.]")[[1]]
  df$Country <- id_list[[1]]
  df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
  df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
  df$model <- id_list[[4]]
  df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
  return(df)  
}

temp_data_CST <- leaf_apply(modelsCST, class=c("plm", "lm"), f = function(model, id){
  return(fittedAndResidData(model, id))
}) %>% do.call(rbind, .)
temp_data_CST$cst <- TRUE

temp_data_noCST <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), f = function(model, id){
  return(fittedAndResidData(model, id))  
}) %>% do.call(rbind, .)
temp_data_noCST$cst <- FALSE

temp_data <- plyr::rbind.fill( temp_data_CST, temp_data_noCST )

# Get the unadjusted, without-energy residuals from models that obey the CST
# and store them in a data frame
r_unadj_noE_CST <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy" & cst==TRUE) %>% 
  select(Year, Country, Residuals, model)
colnames(r_unadj_noE_CST)[colnames(r_unadj_noE_CST)=="Residuals"] <- "r_unadj_noE"
# Merge to get the unadjusted, without-energy values next to each country, year, and model
fitted_and_resid_data <- merge(temp_data, r_unadj_noE_CST)
# Create a column containing ∆r, the difference between a residual value
# and the residual value in the same year for the unadjusted, no-energy case
fitted_and_resid_data[["Dr"]] <- abs(fitted_and_resid_data[["Residuals"]]) - abs(fitted_and_resid_data[["r_unadj_noE"]])

@@

Both historical GDP and fitted GDP are shown in 
Figures~\ref{fig:Graphs CD-fitted}--\ref{fig:Graphs CESekl-fitted}.
%
<<Graphs CD-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Cobb-Douglas models. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CD" & cst==FALSE))
@
%
<<Graphs CESkle-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($kl$)$e$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($le$)$k$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="lek" | nest=="kl")))
@
%
<<Graphs CESekl-fitted, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="CES models with ($ek$)$l$ nesting. Solid line is historical GDP.">>=
fittedGraph(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="ekl" | nest=="kl")))
@

@@

Figures~\ref{fig:Graphs CD-resid}--\ref{fig:Graphs CESekl-resid}
show fitting residuals for all models.
%
<<Graphs CD-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for Cobb-Douglas models.">>=
residualPlot(filter(fitted_and_resid_data, model=="CD" & cst==FALSE))
@
%
<<Graphs CESkle-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(kl$)$e$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="kle" | nest=="kl")))
@
%
<<Graphs CESlek-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with ($le$)$k$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest == "lek" | nest=="kl")))
@
%
<<Graphs CESekl-resid, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Fitting residuals for CES models with $(ek$)$l$ nesting.">>=
residualPlot(filter(fitted_and_resid_data, model=="CES" & cst==FALSE & (nest=="ekl" | nest=="kl")))
@

@@

  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=Dlambda, color=energy), shape=1) +

@@

  geom_point(data=filter(AllCoeffs, cst==TRUE & model=="CD"), mapping=aes(x=Dmse, y=percDlambda, color=energy), shape=1) +

@@

  geom_point(data=filter(AllCoeffs, model=="CES" & cst==FALSE), mapping=aes(x=Dmse, y=Dlambda, color=energy, shape=nest)) +
  geom_point(data=filter(AllCoeffs, model=="CES" & cst==TRUE), mapping=aes(x=Dmse, y=Dlambda, color=energy), shape=1) +

@@

  geom_point(data=filter(AllCoeffs, model=="CES" & cst==FALSE), mapping=aes(x=Dmse, y=percDlambda, color=energy, shape=nest)) +
  geom_point(data=filter(AllCoeffs, model=="CES" & cst==TRUE), mapping=aes(x=Dmse, y=percDlambda, color=energy), shape=1) + 

@@

@@

%' %++++++++++++++++++++++++++++++
%' \subsection{Dynamic Solow residuals} 
%' \label{sec:dynamic_solow_residuals}
%' %++++++++++++++++++++++++++++++
%' 
%' <<Dynamic Solow Residuals>>=
%' # Calculate finite difference approximations of time derivatives for output and all factors of production
%' dynamicSR <- function(model, id, lag=1, pad="tail"){
%'   id_list <- strsplit(id, split = "[.]")[[1]]
%'   energy <- id_list[[3]]
%'   modelStr <- id_list[[4]]
%'   
%'   f <- model$formula
%'   denom <- summands(f, n=length(summands(f))) # time variable
%'   nums <- all.vars(f) [1:(length(all.vars(f))-1)] # y, k, l, and e (if present)
%'   df <- getData(model)[c(denom, nums)] # columns are: iYear, y, k, l, and e (if present)
%'   
%'   # Calculate names for the columns of finite difference derivatives
%'   derivNames <- lapply(1:length(nums), function(i){paste0("d", nums[[i]], ".d", denom)}) %>% unlist
%'   derivCols <- lapply(1:length(nums), function(i){
%'     # Calculate finite difference derivative columns
%'     ediff(df[[i+1]], pad=pad, lag=lag) / ediff(df[[1]], pad=pad, lag=lag)
%'   }) %>% do.call(cbind, .) %>% as.data.frame
%'   names(derivCols) <- derivNames  
%'   df <- cbind(derivCols, df[ , c(nums, denom)]) # puts derivatives first, then raw numbers, then time
%'   df[["dydtOvery"]] <- 1 / df[[ nums[[1]] ]] * df[[ derivNames[[1]] ]]
%'   coeffs <- naturalCoef(model)
%'   df <- cbind(df, coeffs)
%'   
%'   if (modelStr == "CD"){
%'     lambda_fd <- 
%'       df[["dydtOvery"]] - 
%'       df[["alpha_1"]] / df[[nums[[2]]]] * df[[derivNames[[2]]]] - 
%'       df[["alpha_2"]] / df[[nums[[3]]]] * df[[derivNames[[3]]]]
%'     if (energy == "withE"){
%'       # Subtract the energy term also
%'       lambda_fd <- lambda_fd - 
%'         df[["alpha_3"]] / df[[nums[[4]]]] * df[[derivNames[[4]]]]
%'     }
%'     # Fraction of growth explained by Solow Residual
%'     solowFrac <- lambda_fd / df[["dydtOvery"]]
%'     # Add to data frame
%'     df$lambda_fd <- lambda_fd
%'     df$solowFrac <- solowFrac
%'     return(df)
%'   } else if (modelStr == "CES"){
%'     # Assign nice variable names so that the equations below are easier to read (and debug, and maintain)
%'     # This code is nest-aware.
%'     n <- model$nest
%'     for (i in 1:length(model$nest)){
%'       df[[paste0("x", i)]] <- df[[ nums[ model$nest[i]+1 ] ]] # GDP is 1st col, add 1 to get to the factors of production
%'       df[[paste0("dx", i, "dt")]] <- df[[ derivNames[ model$nest[i]+1 ] ]]
%'     }
%'     if (energy == "noE"){
%'       # Calculate terms for the finite-difference Solow residual (lambda_fd)
%'       df <- df %>% 
%'         transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
%'         transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
%'         transform(lambda_fd = dydtOvery + 1/b * dbdtOverrho1) %>%
%'         transform(solowFrac = lambda_fd / dydtOvery)
%'       return(df)
%'     } else if (energy == "withE"){
%'       df <- df %>% 
%'         transform(b = delta_1 * x1^(-rho_1) + (1 - delta_1) * x2^(-rho_1)) %>%
%'         transform(a = delta * b^(rho/rho_1) + (1 - delta) * x3^(-rho)) %>%
%'         transform(dbdtOverrho1 = -delta_1 * x1^(-(rho_1+1))*dx1dt - (1 - delta_1) * x2^(-(rho_1+1))*dx2dt) %>%
%'         transform(dadtOverrho = delta * b^(rho/rho_1 - 1) * dbdtOverrho1 - (1 - delta) * x3^(-(rho+1))*dx3dt) %>%
%'         transform(lambda_fd = dydtOvery + 1/a * dadtOverrho) %>%
%'         transform(solowFrac = lambda_fd / dydtOvery)
%'       return(df)
%'     } else {
%'       stop (paste("Unknown energy:", energy))
%'     }
%'   } else {
%'     stop(paste("Unknown modelStr", modelStr))
%'   }
%' }
%' 
%' lag.to.use <- 1 # years
%' temp_data <- leaf_apply(models, class=c("CDEmodel", "cesModel", "LINEXmodel"), 
%'                         f = function(model, id, lag=lag.to.use, pad="tail"){
%'   df <- dynamicSR(model, id=id, lag=lag, pad=pad)
%'   id_list <- strsplit(id, split = "[.]")[[1]]
%'   df$Country <- id_list[[1]]
%'   df$flavor <- if (id_list[[2]] == "unadjusted") "Unadjusted" else "Quality-adjusted"
%'   df$energy <- if (id_list[[3]] == "noE") "Without energy" else "With energy"
%'   df$model <- id_list[[4]]
%'   df$nest <- if (length(id_list) > 4) id_list[[5]] else NA
%'   df$Year <- getData(model)[["Year"]]
%'   return(df)  
%' }) %>% do.call(plyr::rbind.fill, .)
%' 
%' # Get the unadjusted, without-energy Solow residuals and store them in a data frame
%' sr_unadj_noE <- filter(temp_data, flavor=="Unadjusted" & energy=="Without energy") %>% 
%'                 select(Year, Country, lambda_fd, model)
%' colnames(sr_unadj_noE)[colnames(sr_unadj_noE)=="lambda_fd"] <- "sr_unadj_noE"
%' # Merge to get the unadjusted, without-energy values next to each country, year, and model
%' solowResiduals <- merge(temp_data, sr_unadj_noE)
%' 
%' # Create a column containing ∆lambda_df, the difference between a lambda_df value
%' # and the lambda_df value in the same year for the unadjusted, no-energy case
%' solowResiduals[["Dlambda_fd"]] <- abs(solowResiduals[["lambda_fd"]]) - abs(solowResiduals[["sr_unadj_noE"]])
%' 
%' # Relevel factors before plotting
%' solowResiduals$Country <- relevelFactor(as.factor(solowResiduals$Country), c("PT", "UK"))
%' solowResiduals$flavor <- relevelFactor(as.factor(solowResiduals$flavor), c("Unadjusted", "Quality-adjusted"))
%' solowResiduals$energy <- relevelFactor(as.factor(solowResiduals$energy), c("Without energy", "With energy"))
%' solowResiduals$nest <- relevelFactor(as.factor(solowResiduals$nest), c("kl", "kle", "lek", "ekl"))
%' 
%' # Make data viewing easier
%' solowResiduals <- arrange(solowResiduals, Country, model, flavor, energy, nest)
%' @
%' 
%' <<DynamicSR_Graph>>=
%' dynamicSRGraph <- function(data){
%' ggplot(data=data) +
%'   geom_line(data=filter(data, energy=="Without energy"), mapping=aes(x=Year, y=dydtOvery)) +
%'   geom_line(mapping=aes(x=Year, y=lambda_fd, linetype=energy)) +
%'   scale_linetype_manual(values=c(2,3), name="") +
%'   facet_grid(Country ~ flavor) +
%'   scale_x_continuous(breaks=c(1960, 1980, 2000)) +
%'   xlab("") + 
%'   ylab(expression(lambda[ij])) +
%'   xy_theme()  
%' }
%' @
%' 
%' ``Dynamic'' Solow residuals~($\lambda_{i,j}$) can be calculated 
%' for the Cobb-Douglas and CES models
%' as shown in Appendices~\ref{sec:dynamic_sr_CD} and~\ref{sec:dynamic_sr_CES},
%' respectively.
%' 
%' Figure~\ref{fig:DynamicSR_CD_Graph} shows the evolution 
%' of the Solow residual~($\lambda_{i,j}$) over time
%' as calculated by Equation~\ref{eq:lambdaij} 
%' with the timespan for finite differences ($t_j - t_i$) set to 
%' \Sexpr{lag.to.use}~\Sexpr{if (lag.to.use == 1) "year" else "years"}.
%' %
%' <<DynamicSR_CD_Graph, fig.pos="H", fig.align='center', fig.width=6.5, fig.height=4, fig.cap="Dynamic Solow residuals for the Cobb-Douglas model. Dashed lines are $\\lambda$ values. Solid lines are $\\frac{1}{y}\\frac{\\mathrm{d} y}{\\mathrm{d} t}$.">>=
